Does Anybody Really Know What Time It Is?\\ Automating the Extraction of Date Scalars

### 2016-04-30 cikm2016/nlp.tex
More text trimming

@@

The parsed output is then converted to the ICU Date Format Language described in Section IIA in order to effectively cross-validate results from the MDL algorithm. While this grammar accounts for the use of meridian markers `a.m.' and `p.m.' for the 12-hour format, ICU does not support these tokens, and they are simply ignored.

@@

\item \textbf{LeapYear Restriction for \texttt{February}}:
While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for \texttt{February}, based on whether the year is leap year or not. For this, we define a new symbol in the grammar called \texttt{LeapYear}. If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$.
\end{itemize}

However, Equation $4$ is just an approximation. The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$. Equations $3$ and $4$ are now added as constraints to the \texttt{Days} symbol in the grammar.

@@

The production rules are of the form $ X \rightarrow \beta$ where $X$ is the left-hand side of the rule, while $\beta$ is the right-hand side.

Given a CFG grammar $G$, we define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.  For any $X \rightarrow \beta \in \tau(G)$, the probability $p(X \rightarrow \beta) \ge 1$. In addition, $\sum_{(X \rightarrow \beta) \in \tau_{G}} = 1$. The parser then chooses the parse tree with the maximum probability, \textit{i.e.} given a date-time string $s$, as input, we determine the highest scoring parse tree $X \rightarrow \beta$ as  $max_{(X \rightarrow \beta) \in \tau(s)} p(X \rightarrow \beta)$.

Assigning initial probability weights to the grammar for computing the parse with the highest weight, tends to be a more scalable and flexible approach to disambiguation rather than building a deep knowledge ontology. We bootstrap the probability weights in the \texttt{PCFG}, by setting initial probability weights to the grammar rules as follows:

\begin{itemize}
\item Similar to the MDL ranking properties described in Section IIIG, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}. In practice, assigning $p(\texttt{DateGrammar} \rightarrow \beta_{1}) = 0.9$ and $p(\texttt{TimeGrammar} \rightarrow \beta_{2}) = 0.7$ leads to optimal ranking of the parse trees.

@@

Having defined initial probabilistic weights to the \texttt{PCFG}, we employ supervised learning with a known training set of date time formats to estimate the rule probabilities.  The training corpus is a set of files obtained from an online data analysis corpus, described in Section 6.

@@

In addition to the columnar data context, the locale is also used to help maximize the probabilistic likelihood of a particular date-time pattern. The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors. However, similar to the locale sensitivity issue described in Section IIIF, the locale and the data may not always match. The dominant pattern is simply computed over the data column based on the ranked set of parse tree results.

>>>>

\vidya{need to add a short intro here}

@@

The parsed output is then converted to the ICU Date Format Language in order to effectively cross-validate results from the MDL algorithm. While this grammar accounts for the use of meridian markers `a.m.' and `p.m.' for the 12-hour format, ICU does not support these tokens, and they are simply ignored.

@@

\item \textbf{LeapYear Restriction for \texttt{February}}:
While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for \texttt{February}, based on whether the year is leap year or not. For this, we define a new symbol in the grammar called \texttt{LeapYear}. If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$. However, this is just an approximation. The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$.

@@

Given a CFG grammar $G$, the production rules are of the form $ X \rightarrow \beta$ where $X$ is the left-hand side of the rule, while $\beta$ is the right-hand side. We define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.  For any $X \rightarrow \beta \in \tau(G)$, the probability $p(X \rightarrow \beta) \ge 1$. In addition, $\sum_{(X \rightarrow \beta) \in \tau_{G}} = 1$. The parser then chooses the parse tree with the maximum probability, \textit{i.e.} given a date-time string $s$, as input, we determine the highest scoring parse tree $X \rightarrow \beta$ as  $max_{(X \rightarrow \beta) \in \tau(s)} p(X \rightarrow \beta)$.

Assigning initial probability weights to the grammar for computing the parse with the highest weight, tends to be a more scalable and flexible approach to disambiguation rather than building a deep knowledge ontology. We bootstrap the probability weights in the \texttt{PCFG} as follows:

\begin{itemize}
\setlength\itemsep{0em}
\item Similar to the MDL ranking properties, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}. In practice, assigning $p(\texttt{DateGrammar} \rightarrow \beta_{1}) = 0.9$ and $p(\texttt{TimeGrammar} \rightarrow \beta_{2}) = 0.7$ leads to optimal ranking of the parse trees.

@@

Having defined initial probabilistic weights to the \texttt{PCFG}, we employ supervised learning with a known training set of date time formats to estimate the rule probabilities.  The training corpus is a set of files obtained from an online data analysis corpus, described in Section 5.

@@

In addition to the columnar data context, the locale is also used to help maximize the probabilistic likelihood of a particular date-time pattern. The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors. However, similar to the locale sensitivity issue described in Section 3.6, the locale and the data may not always match. The dominant pattern is simply computed over the data column based on the ranked set of parse tree results.

### 2016-04-30 cikm2016/testing.tex
More tightening
More text trimming

\subsubsection{NULL Filtering}
Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of \texttt{NULL}:
\begin{itemize}
\setlength\itemsep{0em}
\item Values containing the substring \texttt{NULL}
\item Values containing no digits and at most one non-whitespace character (\eg \texttt{" / / "})
\item Values on a list of empirically determined common \texttt{NULL} values (\eg \texttt{0000-00-00}, \texttt{NaN})
\end{itemize}
%Columns that had no remaining valid samples were discarded.

@@

Once the training data was analyzed, it was grouped by date format. A sample of each produced date format was manually labeled. This allowed us to quickly skip over very common formats like \texttt{MM/dd/yyyy} and focus our efforts on much less common formats. The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.

Further manual tagging focused on the files most likely to represent dates. Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date''). A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.

>>>>

\subsubsection{NULL Filtering}
Each column was then reduced to a maximum sample set of 32. We excluded domain values that either contain the substring \texttt{NULL}, contain no digits and at most one non-whitespace character (\eg \texttt{" / / "}), or any empirically determined common \texttt{NULL} value (\eg \texttt{0000-00-00}, \texttt{NaN}).
%Columns that had no remaining valid samples were discarded.

@@

Once the training data was analyzed, it was grouped by date format. A sample of each produced date format was manually labeled. This allowed us to quickly skip over very common formats like \texttt{MM/dd/yyyy} and focus our efforts on much less common formats. The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.
\vidya{I don't understand this para below. Can be abridged and tightened with the previous para}
Further manual tagging focused on the files most likely to represent dates. Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date''). A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.
